---
title: "Prediction Assignment"
author: "Justin Meyer"
date: "January 18, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE, comment = NA)
```

## Building the Model

First I read in the csv files. The data come from http://web.archive.org/web/20161224072740/http:/groupware.les.inf.puc-rio.br/har.

```{r}

testing <- read.csv("pml-testing.csv")
training <- read.csv("pml-training.csv")

```

Then I ran some descriptive statistics on the data. I chose these fields because they weren't missing any data and because they seemed the most relevant to the outcome that I want to predict.

```{r}

training <- subset(training, 
                   select = c("classe", 
            "roll_belt", "pitch_belt", "yaw_belt", "total_accel_belt",
            "roll_arm", "pitch_arm", "yaw_arm", "total_accel_arm",
            "roll_dumbbell", "pitch_dumbbell", "yaw_dumbbell", "total_accel_dumbbell",
            "roll_forearm", "pitch_forearm", "yaw_forearm", "total_accel_forearm"))

testing <- subset(testing, 
                   select = c( 
            "roll_belt", "pitch_belt", "yaw_belt", "total_accel_belt",
            "roll_arm", "pitch_arm", "yaw_arm", "total_accel_arm",
            "roll_dumbbell", "pitch_dumbbell", "yaw_dumbbell", "total_accel_dumbbell",
            "roll_forearm", "pitch_forearm", "yaw_forearm", "total_accel_forearm"))

summary(training)

```

Next I split the training set into a training set and a validation set since I can't use the provided test set to validate my model since the test set doesn't include the classe field.

```{r}

data <- training
rm(training)

library(caret)
set.seed(73)

trainIndex <- createDataPartition(data$classe, 
                                  p = .8, # Proportion in the index
                                  list = FALSE, # Output type
                                  times = 1) # Number of partitions

training <- data[trainIndex,]
validation  <- data[-trainIndex,]

rm(data)

```

Next I created a conditional inference tree model based on the training data. I chose a conditional inference tree because it works with the specified classification problem, provided accurate results on the training data, and ran quickly. The model predicts the field "classe" using the other available fields.

**How I used cross validation: I used k-fold cross validation to evaluate the accuracy of the model using the training data. This is how the accuracy of the model on the training set is produced. This information suggests whether or not the model will be useful at predicting the outcome variable using new data sets.**

```{r}

# Specify parallel processing
library(parallel)
library(doParallel)
cluster <- makeCluster(detectCores() - 1) # leave out one core
registerDoParallel(cluster)

# Specify how to do cross validation
train_control <- trainControl(method = "cv", 
                              number = 3)

# Fit the model
fit <- train(classe ~ ., 
                data = training, 
                method = "ctree",
                trControl = train_control)

```

# Checking the Model

The below output shows that the model performs well on the training data, with an accuracy of `r round(max(fit$results$Accuracy) * 100, 1)`%.

```{r}

fit

```

The following table shows the importance of the variables to the model.

```{r}

varImp(fit)

```

The following chart shows the relationship between p value threshold and accuracy.

```{r}

plot(fit)

```

## Checking the Out of Sample Error with the Validation Set

Next I used the validation data set to check the accuracy of my model on new data. I made a prediction for each of the records in the validation set and then compared that to the actual "classe" value. **I expect that the out of sample error will be about 12% since accuracy on the validation set is about 88%, with a 95% confidence interval of about 87% to 89%.**

```{r}

validation_prediction <- predict(fit, validation)

confusionMatrix(validation_prediction, validation$classe)

```

## Using the Model to Make Predictions on the Test Set

Finally, I use the ctree model to predict the class for each of the testing set records.

```{r}

predict(fit, testing)

```
